Hijacking Simulators with Universal Probabilistic Programming

### 2019-05-20 paper/neurlps2019/main.tex
updated

Probabilistic programming\cite{gordon2014probabilistic,le2016inference} is a machine learning paradigm that enables users to write, and perform inference in, complex probabilistic models in a programmatic, yet efficient and simplified way. 
The source code of the given model, or equivalently the \emph{program}, is then transformed into a program density, which can
be viewed as the joint density of the user-specified model.
Having direct access to the source code enables probabilistic programming systems~(PPSs)
to conduct several processes in an automated fashion, such as 
the automation of complex inference procedures\cite{zhou2019lf,goodman2012church,wood2014new}.
In addition to this, as models can be directly expressed in program code we can encode other structures to make use of the full structure of a given language,
which enables users to integrate PPSs into their standard model development process, increasing efficiency and reducing model iteration time.

One powerful \emph{class} of models are simulators. 
Simulators arise in many industrial \cite{di2015simulator,judd2009numerically,kass1990rapid} and scientific applications~\cite{smith2008towards,bershteyn2018implementation,gleisberg_event_2008} encode years, if not decades, of research and development.
Thus, Simulators by their very nature are
information rich and naturally describe complicated joint densities.
However, in order to write a simulator in a standard probabilistic programming set-up one would have to re-write a simulator in the language of the PPS, which 
in many cases is not feasible, nor efficient due to both the size 
and complexity of the simulator code base. 
However, what if one could arbitrary connect a simulator, written in any
language and connect it directly to a PPS, without having to re-write
the simulator in the given PPS?

>>>>

% One powerful \emph{class} of models are simulators. 
Simulators are useful tools when conducting experiments in the real system 
is impossible, or impractical and they arise in numerous industrial \cite{di2015simulator,judd2009numerically,kass1990rapid} and scientific applications~\cite{smith2008towards,bershteyn2018implementation,gleisberg_event_2008} . They are rich models and encode years, if not decades, of research and development.
The \emph{model}, or equivalently \emph{program}, specified by the simulator defines a joint density $p(x,y)$ and the inputs into the simulator denote priors, $p(x)$.

 def that are difficult to generate represent a large class of generative models, where the simulator itself represents a joint density
$p(x|y)$  arise in many industrial
Thus, Simulators by their very nature are
information rich and naturally describe complicated joint densities.
However, in order to write a simulator in a standard probabilistic programming set-up one would have to re-write a simulator in the language of the probabilistic programming system (PPS), which 
in many cases is not feasible, nor efficient due to both the size 
and complexity of the simulator code base. 
However, what if one could arbitrary connect a simulator, written in any
language and connect it directly to a PPS, without having to re-write
the simulator in the given PPS?

\paragraph{Probabilistic programming}\cite{gordon2014probabilistic,le2016inference} is a machine learning paradigm that enables users to write, and perform inference in, complex probabilistic models in a programmatic, yet efficient and simplified way. 
The source code of the given model, or equivalently the \emph{program}, is then transformed into a program density, which can
be viewed as the joint density of the user-specified model.
Having direct access to the source code enables PPSs
to conduct several processes in an automated fashion, such as 
the automation of complex inference procedures\cite{zhou2019lf,goodman2012church,wood2014new}.
In addition to this, as models can be directly expressed in program code we can encode other structures to make use of the full structure of a given language,
which enables users to integrate PPSs into their standard model development process, increasing efficiency and reducing model iteration time.
